### 2016-04-19 manuscript.tex
scrub of sec 4

\documentclass[preprint,11pt]{aastex} 
\let\captionbox=\undefined
\usepackage[top=1in, bottom=1in, left=1in, right=1in]{geometry}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{mdwlist}
\usepackage{natbib}
\usepackage{natbibspacing}
\usepackage[font={footnotesize}]{caption}
\usepackage{wrapfig}
\usepackage{sidecap}
\usepackage{tabularx}
\usepackage{enumitem}
\usepackage{url}
\setlist[itemize]{noitemsep, topsep=0pt}
\setlist[enumerate]{noitemsep, topsep=0pt}
\setlength{\bibspacing}{0pt}
\setlength{\parskip}{0pt}
\setlength{\parsep}{0pt}
\setlength{\headsep}{0pt}  
\setlength{\topskip}{0pt}
\setlength{\topmargin}{0pt}
\setlength{\topsep}{6pt}
\setlength{\partopsep}{0pt}
\setlength{\footnotesep}{8pt}
\pagestyle{plain}
\citestyle{aa}
\captionsetup[table]{labelsep=space}
\captionsetup[figure]{labelsep=space}

%\newcommand{\Mycite}[1]{{\bf \cite{#1}}}
%\newcommand{\Mycitet}[1]{{\bf \citet{#1}}}
%\newcommand{\Mycitep}[1]{{\bf \citep{#1}}}
%\newcommand{\Mycitealt}[1]{{\bf \citealt{#1}}}
\newcommand{\Mycite}[1]{\cite{#1}}
\newcommand{\Mycitet}[1]{\citet{#1}}
\newcommand{\Mycitep}[1]{\citep{#1}}
\newcommand{\Mycitealt}[1]{\citealt{#1}}

\newcommand{\kvec}{{\bf k}}
\newcommand{\bvec}{{\bf b}}
\newcommand{\shat}{{\hat s}}
\newcommand{\kpr}{{k_\perp}}
\newcommand{\kvpr}{{\kvec_\perp}}
\newcommand{\AI}{{\langle\tilde A*\tilde I\rangle}}
\newcommand{\AItau}{{\AI(\tau)}}
\newcommand{\hMpci}{{h~{\rm Mpc}^{-1}}}
\newcommand{\inch}{$^{\prime\prime}$}
\newcommand{\foot}{$^{\prime}$}
\renewcommand{\deg}{^\circ}
\newcommand{\integral}{\int\limits}

\newcommand{\simgt}{\stackrel{>}{_{\sim}}}
\newcommand{\compress}{\vspace{-0.25in}}
\newcommand{\captionbaseline}{\renewcommand{\baselinestretch}{0.99}} 
\newcommand{\Caption}[4]{\vspace{#1}\renewcommand{\baselinestretch}{#2}\caption{#4}\vspace{#3}}
\def\kperp{k_{\bot}}
\def\eppsilon{{$\varepsilon$ppsilon}}
\def\kpar{k_{\|}}
\def\k{{\bf k}}
\def\sky{{\theta}}
\def\HI{{H{\small I }}}
\def\HII{{H{\small II }}}
\def\xHI{{x_{\rm\HI}}}
\newcommand{\startsquarepar}{%
    \par\begingroup \parfillskip 0pt \relax}
\newcommand{\stopsquarepar}{%
    \par\endgroup}

%%% UNCOMMENTED
\usepackage{subfig}
%\usepackage[countmax]{subfloat}
\title{Hydrogen Epoch of Reionization Array (HERA)}
\author{A.U.Thor}

\begin{document}
\maketitle

\begin{center}
\textbf{Abstract}
\end{center}

The Hydrogen Epoch of Reionization Array (HERA http://reionization.org) is a staged experiment that uses the unique properties of the 21-cm line from neutral hydrogen to probe the Epoch of Reionization (EOR). During this epoch, roughly 0.3 - 1 billion years after the Big Bang, the first galaxies and black holes heated and reionized the early Universe. Direct observation of the large scale structure of reionization and its evolution with time will have a profound impact on our understanding of the birth of the first galaxies and black holes, their influence on the intergalactic medium (IGM), and cosmology.  This paper will provide an overview of the project and describe the design of the system.

\section{Introduction}
The Hydrogen Epoch of Reionization Array (HERA) is a staged experiment to use the redshifted 21\,cm line
of neutral hydrogen to characterize our Cosmic Dawn, from
the formation of the first stars and black holes $\sim0.1~$Gyr after the Big
Bang ($z\sim30$) through the full reionization of the intergalactic medium
(IGM) $\sim1~$Gyr later ($z\sim6$).  By directly observing the large scale
structure of the primordial IGM as it is heated and reionized, HERA complements
other probes, adding transformative capabilities for
understanding the astrophysics and fundamental cosmology of our early universe.
Taking advantage of a new understanding of bright foreground systematics, HERA's
purpose-built radio
interferometer is optimized to deliver high signal-to-noise measurements of
redshifted 21\,cm emission to detect and characterize the Epoch of Reionization (EOR).

Because of these key science questions that may be probed by highly redshifted hydrogen, there has been a resurgence in low frequency arrays  -- most with a primary
objective of measuring the EOR.  These include the Precision Array Probing the Epoch of Reionization (PAPER; \citealt{parsons_et_al2010}), the Murchison Widefield Array (MWA; \citealt{tingay_et_al2013}), the LOw Frequency ARray (LOFAR; \citealt{2013A&A...556A...2V}), and the Long Wavelength Array (LWA; \citealt{2009IEEEP..97.1421E}), as well as systems for existing dish arrays like the Very Large Array (VLA; \citealt{2013AAS...22132804K}) and the Giant Metrewave Radio Telescope (GMRT; \citealt{paciga_et_al2011}).
With the first generation 21\,cm telescopes approaching the limits of their sensitivity however, HERA continues the experiment by bringing significantly more sensitivity to bear
while continuing to optimize its design along the delay-spectrum approach to detect the power spectrum of the EoR (discussed in Sec. \ref{sec:delayapproach}).
At the same time, HERA continues the development of new techniques to extend and improve its capabilities.
Examples include improving foreground subtraction to expand the information content 
in power spectral and imaging measurements and
extending HERA's performance at lower frequencies to probe IGM heating in the pre-reionization epoch (\S\ref{sec:EoX}).

\begin{figure}[h!]
	\centering
	\includegraphics[height=2.0in]{plots/hera_render_logo.pdf}
	\includegraphics[height=2.0in]{plots/hera19H.png}
	\caption{Rendering of the 320-element core (left) of the full HERA-350 array and picture of 19 HERA 14-m, zenith-pointing dishes (with PAPER elements in the background) currently deployed in South Africa (right).} 
	\label{fig:HERApictures}
\end{figure}

The HERA experiment will comprise 
350 14-m parabolic dishes (320 in a dense core $+$ 30 outriggers) in the South African Karoo Radio
Astronomy Reserve (see Fig. \ref{fig:HERApictures} (top) for a rendering).
HERA has been funded under the US National Science Foundation's Mid-Scale Innovations Program to begin building elements optimized for robust power spectrum detection in a 
staged deployment.  The key is to produce inexpensive collecting area optimized to detect the EOR.  
The first 19-element array is complete and, with additional funding from Cambridge University, construction underway on the next 18 elements, for a total of 37.
The South African National Research Foundation SKA-SA group (SKA-SA) is a key partner in the construction and science.

The current 19 elements are testing the HERA element design for performance and manufacturability on location.  The 37 elements, to be completed in 2016, will provide nearly an order of magnitude more 
sensitivity than current instruments and provide a significant chance at detecting the EoR power spectrum signal.  The next proposed phase will build out to 128 elements and use the existing 128 dual-polarization analog and digital signal paths that have been in use with PAPER \citep{ali_et_al2015}, which should provide a very robust detection of the EOR signal and allow some characterization.  Finally, extending to 2020, HERA will build out to 350 elements to further EOR science as a function 
of redshift and spatial scale, potentially producing the first images of the EOR.

HERA is a collaboration of the following Partner institutions:  
Arizona State University (Tempe, AZ USA), 
Brown University (Providence, RI USA), 
University of California Berkeley (Berkeley, CA USA), 
University of California Los Angeles (Los Angeles, CA USA), 
University of Cambridge (Cambridge, UK), 
Massachusetts Institute of Technology (Cambridge, MA USA), 
National Radio Astronomy Observatory (Charlottesville, VA USA), 
University of Pennsylvania (Philadelphia, PA USA), 
Scuola Normale Superiore di Pisa (Pisa, Italy),
SKA-South Africa (Cape Town, South Africa), and
University of Washington (Seattle, WA USA).  
Additional Collaborators are at 
Harvard University (Cambridge, MA USA),  
University of KwaZulu Natal (Durban, South Africa), 
University of Western Cape (Cape Town, South Africa), 
Imperial College London  (London, UK) and 
California State Polytechnic University (Pomona, CA USA).

This paper has the following structure.  Section \ref{sec:science} will give a high-level overview of the science motivating HERA's construction, including primary and secondary science objectives.  Key science results include
measuring the 21\,cm power spectrum with high significance (\S\ref{sec:EoRPowerSpectra});
characterizing parametrized models of reionization and the first ionizing sources (\S\ref{sec:EoRPowerSpectra});
improving cosmic microwave background (CMB) constraints on primordial density fluctuations and neutrino masses by removing the optical depth degeneracy (\S\ref{sec:tau}); and
directly imaging large scale reionization structure (\S\ref{sec:imaging}).
Additionally, beyond these directly supported applications, HERA's public data products and instrumentation platform enable other high-impact science, including 
21\,cm cross-correlation (e.g. with JWST, WFIRST, CO, CII, and Ly$-\alpha$; \S\ref{sec:crossCorr});
searching for auroral radio bursts from exoplanetary magnetospheres (\S\ref{sec:exoplanets}); and
performing triggered, low-frequency follow up and characterization of fast radio bursts (\S\ref{sec:FRBs}).

Section \ref{sec:eormeas} will present the techniques used to make the measurement.  Section \ref{sec:requirements} will present the high-level requirements and Section \ref{sec:design} will present the system description..   Section \ref{sec:status} will present a brief status and outline the deployement timetable and Section \ref{sec:conclusion} will conclude by summarizing and providing additional context.

\section{Scientific Background}
\label{sec:science}

The {\it Cosmic Dawn} of our universe is one of the last unexplored
frontiers in cosmic history.  This history is summarized in Fig. \ref{fig:cosmos}, starting with the Big Bang on the left. The hot, young universe expands and cools slowly in the background while gravitational instability around concentrations of dark matter causes primordial density fluctuations to grow.
The Cosmic Dawn represents a specific epoch in this growth, where the first stars and galaxies formed and illuminated the Universe en route to forming the astronomical structures we see today. Ultimately, this early population of gravitationally condensed material produced sufficiently energetic flux to reionize the intergalactic medium (IGM) from its previous neutral state in a period called the Epoch of Reionization.  This period is shown in Fig \ref{fig:cosmos} as the rapid transition from separated large reionized ``bubbles'' to the merged reionized state and to structures that begin to resemble the denizens of our current universe. The structure of the 
IGM thus contains a panoply of information about the underlying astrophysical and cosmological phenomena governing cosmic evolution.

The evolution of the cosmic structure depends on the local and average cosmic density, the relative velocities of
baryons and dark matter, and the sizes and clustering of the first galaxies to form.
But it also depends on the constituents of those first galaxies---so-called Population
III stars (stars formed very early with little to no elements heavier than helium), later generation stars, stellar remnants, X-ray binaries,
and early supermassive black holes.  Bulk properties like
ultraviolet and X-ray luminosities and spectra also affect the thermal and
ionization states of the IGM.  The wealth of unexplored physics during the Cosmic Dawn,
culminating in the Epoch of Reionization, 
led the most recent US National Academies astronomy decadal survey entitled {\em New Worlds, New Horizons} to highlight it as one of the top three ``priority science objectives'' for
the decade \citep{NWNH}.

\begin{figure}[h!]
	\centering
	\includegraphics[width=0.7\textwidth]{plots/cosmicEvo.png}
	\caption{Rendering of cosmic evolution from just after the Big Bang to today (background image credit Loeb/{\em Scientific American}).  The blue line shows the frequency of the red-shifted hydrogen line (rest frequency 1420 MHz) as a function of redshift (lower horizontal axis) and cosmic time (upper horizontal axis).  The white dashed horizontal lines are the HERA EoR band and the blue are the extended frequency goal.  Determining where the structure in the background really falls is the
goal of HERA.} 
	\label{fig:cosmos}
\end{figure}

Exploring the interplay of galaxies and large-scale structure during the EOR
requires complementary observational approaches. Measurements of the Cosmic Microwave Background 
(CMB; the escaping radiation permeating the universe after the universe became
transparent to its radiation by the recombination of the protons and electrons about 400,000 years after the Big Bang)  by COBE, WMAP and Planck provide initial conditions for structure formation.  Thomson
scattering of CMB photons by the ionized particles constrains the integrated column of 
ionized gas and kinetic Sunyaev-Zel'dovich measurements constrain the power spectrum of cosmic structures.
But even with these measurements, the detailed evolution of the IGM 
is only loosely constrained \citep{haiman_holder2003,mortonson_hu2008,zahn_etal2012,mesinger_et_al2012}.
Lyman-$\alpha$ absorption features in quasar and $\gamma$-ray burst spectra give 
ionization constraints at the tail end of reionization 
($z\la7$, \citealt{fan_et_al2006, mcgreer_et_al2015}), but these features 
saturate at low neutral fractions
($x_{\rm HI} \la 10^{-4}$), where $x_{\rm HI}$ is the ratio of neutral hydrogen to the total baryonic density.

Measurements of galaxy populations in
deep {\it Hubble Space Telescope} observations 
have pinned down the bright end of the galaxy luminosity function at $z \la 8$
\citep{schenker_et_al2013, bouwens_et_al2015} and are pushing deeper
(e.g.~\citealt{mcleod_et_al2015}), but producing a consistent ionization history
requires broad extrapolations to
lower-mass galaxies and ad hoc assumptions about the escape fraction of
ionizing photons and the faint-end cutoff of ionizing galaxies
\citep{robertson_et_al2015, bouwens_et_al2015_reion}. 
Similarly, deducing the ionization state of the IGM from quasar proximity zones
\citep{carilli_et_al2010, bolton_et_al2011, bosman_becker2015} and the
demographics of Ly-$\alpha$ emitting galaxies \citep{fontana_et_al2010,
schenker_et_al2012, treu_et_al2012, dijkstra_et_al2014} is uncertain and
highly model-dependent.  See Fig.~\ref{fig:IonHist} for these constraints on the hydrogen neutral fraction as a function of redshift.
As shown, existing probes are limited in
their ability to constrain reionization, and will be for the foreseeable future.
HERA will use
another complementary probe --- the 21\,cm ``spin-flip" transition of neutral hydrogen
--- to bring transformative new capabilities in this area.  The next sub-sections outline these goals.

\begin{figure}[h!]
\centering
    \includegraphics[width=0.51\textwidth,clip]{plots/ionHist.pdf}
  \caption{Combining direct constraints on $x_\text{HI}$, the hydrogen neutral fraction, as a function of redshift (black points) with \emph{Planck} priors yields
an inferred $95\%$ confidence region (gray).  HERA constraints with (dark red) and without (pale red) 
a conservative $25\%$ modeling error in the $21\,\textrm{cm}$ power spectrum can dramatically narrow this confidence region.}
	\label{fig:IonHist}
\end{figure}

\subsection{Precision Constraints on Reionization}
\label{sec:EoRPowerSpectra}

HERA's primary science goal is to transform our understanding of the first stars, galaxies, 
and black holes, and their role in driving reionization. 
Through power-spectral measurements of the 21\,cm line of hydrogen in the primordial IGM,
HERA will be able to direct constrain the topology and evolution of reionization, 
opening a unique window into the complex astrophysical interplay between the 
first luminous objects and their environments.
The spectral nature of 21\,cm cosmology means that 
the signal at each observing frequency can be associated with an emission time (or distance) to determine both the time evolution
and three-dimensional spatial structure of ionization in the IGM.
This 3D structure encodes information about the clustering properties of galaxies,
allowing us to distinguish between models, even if they predict the same ionized fraction. 
With a new telescope optimized for 3D power-spectral measurements and with support for theoretical
modeling efforts, the HERA program will advance our understanding of early galaxy formation and cosmic reionization.

\begin{figure}[h!]
\centering
    \includegraphics[width=0.49\textwidth,clip]{plots/sensitivity_wideband.pdf}
  \caption{ 1$\sigma$ thermal noise errors on $\Delta^2(k)$, the 21\,cm power spectrum, at $k\!=\!0.2$\,$h$\,Mpc$^{-1}$ (the dominant error at that $k$)
with 1080 hours of integration (black)
compared with various heating and reionization models (colored).  Note that only predicted results using the delay-spectrum approach are shown.}
	\label{fig:Sensitivities}
\end{figure}

HERA builds on the advances of first-generation
21\,cm EoR experiments led by HERA team members, particularly 
the Donald C. Backer Precision Array for Probing the Epoch of Reionization (PAPER; \citealt{parsons_et_al2010}),
the Murchison Widefield Array (MWA; \citealt{bowman_et_al2012,tingay_et_al2013}),
the MIT EoR experiment (MITEoR; \citealt{zheng_et_al2014}) and the Experiment to Detect the Global EoR Step (EDGES; \citealt{bowman_rogers2010}).
Recent measurements
have produced the first astrophysically constraining upper limits on the 21\,cm EoR power spectrum, 
providing evidence for significant heating in the IGM prior to reionization \citep{parsons_etal2014, parsons_et_al2015, ali_et_al2015,pober_et_al2015}.
However, current experiments cannot expect more than marginal detections of the EoR signal. 
Figure~\ref{fig:Sensitivities} compares telescope sensitivities as a function of redshift to models of 
the evolving, spherically averaged 21\,cm EoR power spectrum, characterized by the dimensionless power spectrum parameter $\Delta^2 (k) \equiv k^3 P(k) / 2 \pi^2$.
The expected performance of HERA relative to current and planned telescopes to detect the peak of reionization (as well as the total collecting area is shown in 
Table~\ref{tab:signif}.

%In contrast to current experiments,
%HERA's optimized design and collecting area make it capable of
%high-significance detections of virtually any realistic reionization scenario, even using only currently demonstrated foreground
%mitigation strategies (see Table~\ref{tab:signif}). The development of enhanced foreground modeling and subtraction further improve the achievable results.

\begin{SCtable}
\caption{\hspace{-1.2mm}: Predicted SNRs of 21\,cm experiments for an EoR model with 50\% ionization at $z=9.5$, with 1080 hours observation, integrated over a $\Delta z$ of $0.8$.
Foreground avoidance represents an analysis comparable to \cite{ali_et_al2015}, whereas foreground modeling allows signifcantly more $k$ modes of the cosmological signal to be recovered.}
\small
 \centering
 \begin{tabular}{c||r||r|r} 
%\begin{deluxetable}{c||r||r|r}
%\tabletypesize{\small}
%\tablecaption{\small
\hline
%\startdata
Instrument & \shortstack{Collecting \\ Area (m$^2$)} & \shortstack{Foreground \\Avoidance} & \shortstack{Foreground \\Modeling} \\
\hline
PAPER & 1,188 & 0.77$\sigma$ & 3.04$\sigma$ \\
MWA & 3,584 & 0.31$\sigma$ & 1.63$\sigma$ \\
LOFAR NL Core & 35,762 & 0.38$\sigma$ & 5.36$\sigma$ \\
%\textbf{HERA-127} & \textbf{19,500} & \textbf{10.88} & \textbf{35.65} \\
\textbf{HERA-350} & \textbf{53,878} & \textbf{23.34$\sigma$} & \textbf{90.97}$\boldsymbol{\sigma}$ \\
SKA1 Low Core & 416,595 & 13.4$\sigma$ & 109.90$\sigma$
%\enddata
\end{tabular}
%\Caption{-0.1in}{0.99}{-0.1in}
\hspace{-0.1in}
\label{tab:signif}
\end{SCtable}

\begin{figure}[h!]
\centering
    \includegraphics[width=0.56\textwidth,clip]{plots/LikelihoodContours_smaller_avoid_All3_no_table}
  \caption{Projected likelihood contours from an MCMC analysis for astrophysical parameters of reionization. Model parameters are $T_\textrm{vir}^\textrm{min}$ (minimum virial temperature of ionizing galaxies); $R_\textrm{mfp}$ (mean free path of ionizing photons in HII regions); and $\zeta_0$ (ionizing efficiency of galaxies).  Also shown are constraints on the derived ionizing escape fraction, $f_\textrm{esc}$. }
	\label{fig:paramConstraints}
\end{figure}

HERA's 21\,cm measurements can be used in conjunction with our semi-analytic models to constrain the ionization history. 
The red band in Fig.~\ref{fig:IonHist} shows the forecasted 95\% confidence region derived from HERA data after marginalizing over astrophysical and cosmological parameters.
Figure~\ref{fig:paramConstraints} shows the results of a Markov Chain Monte
Carlo (MCMC) pipeline for fitting models to 21\,cm power spectrum data \citep{greig_and_mesinger2015}, which we have conservatively limited to the range $8 < z < 10$ (although in practice a much broader bandwidth will be available; see Sec. \ref{sec:freqs}).
Based on the excursion-set formalism of
\citet{furlanetto_et_al2004} and the 21cmFAST code \citep{mesinger_et_al2011},
this code models the astrophysics of
reionization with three free parameters (see Figure~\ref{fig:paramConstraints} for details). 
While the existing experiment with the most collecting area, 
the LOw Frequency ARray (LOFAR; \citealt{2013A&A...556A...2V, yatawatta_et_al2013}),
provides some ability to constrain these parameters,
HERA's constraints are significantly more precise and are comparable
to what could be achieved with the SKA. 
Additionally, HERA's constraints enable principal component parameterizations of the
sky-averaged $21\,\textrm{cm}$ signal measurements pursued by experiments such as EDGES, increasing their signal-to-noise and thus their science return \citep{liu_parsons2015}.

\subsection{Secondary Scientific Objectives}%\\Precision Cosmology, Reionization Imaging, and High-$z$ IGM X-Ray Heating Science}

\emph{\textbf{Precision Cosmology.}}
\label{sec:tau}
By advancing our understanding of reionization astrophysics, HERA will improve CMB constraints on 
fundamental cosmological parameters by
removing the optical depth
$\tau$ as a ``nuisance'' parameter. HERA breaks the degeneracy between the
constraints on $\tau$ and the sum of the neutrino masses $\sum m_\nu$, which has
been identified as a potential problem for Stage 4 CMB lensing experiments. A
HERA-informed estimate of $\tau$ enables CMB lensing experiments to achieve a
$0.012\,\textrm{eV}$ error on $\sum m_\nu$ \citep{liu_et_al2015}. This would
represent a $\sim$5$\sigma$ cosmological detection of the neutrino masses even
under the most pessimistic assumptions
% that their sum is the minimum value of $0.058\,\textrm{eV}$ 
still allowed by neutrino oscillation experiments
\citep{allison_et_al2015}, making HERA key to our understanding of neutrino
physics. HERA's estimate of $\tau$ would also break the degeneracy between
$\tau$ and the amplitude of matter fluctuations (expressed in Fig.
\ref{fig:sigma8Tau} as $\sigma_8$) that arises when using only
CMB data. HERA effectively reduces error bars on $\sigma_8$ by more
than a factor of three \citep{liu_et_al2015}, potentially elucidating
current tensions between cluster cosmology constraints and those from primary
CMB anisotropies.

\begin{figure}[h!]
\centering
    \includegraphics[width=0.50\textwidth,clip]{plots/sigmaTau.pdf}
\caption{Likelihood contours ($68\%$ and $95\%$) for $\sigma_8$ and $\tau$ using \emph{Planck} constraints (blue) and 
adding HERA data (red). 
%Light and heavy contours signify $68\%$ and $95\%$ confidence regions, respectively. 
The $21\,\textrm{cm}$ constraints break the CMB degeneracy between the amplitude of density fluctuations and the optical depth, improving constraints on both.}
	\label{fig:sigma8Tau}
\end{figure}

\emph{\textbf{First Images of the Reionization Epoch.}}
\label{sec:imaging}
In addition to measuring the power spectrum, there is the potential for HERA  
 to directly image the IGM during reionization
over the 1440 deg$^2$ stripe that transits overhead, which is comparable to future WFIRST
large area near-IR surveys.
 After 100 hours on a single field (achievable in 200 nights), HERA reaches a surface brightness 
sensitivity of 50 $\mu$Jy/beam (synthesized beam FWHM $\sim$ 24$'$) compared 
to the brightness temperature fluctuations of up to 400 $\mu$Jy/beam in typical 
reionization models (see Fig.~\ref{fig:LSS}). From the standpoint of sensitivity alone, HERA is capable of 
detecting the brightest structures at $z=8$ with SNR $>$ 10. Additionally, the design 
of HERA places it in a unique position to directly explore calibration techniques
(e.g. redundant calibration, \citealt{liu_et_al2010,zheng_et_al2014}), while retaining a high quality 
point spread functions for imaging and identifying foregrounds.

%\subsubsection{The 21\,cm Signal in the Pre-Reionization Epoch}

%\begin{wrapfigure}{R}{0.5\textwidth}
\begin{figure}[b!]
\centering
\vspace{-15pt}
    %\includegraphics[width=.5\textwidth,clip]{plots/imaging/HERA_z8_SNR_annotated_2015.png}
    \includegraphics[width=1.0\textwidth,clip]{plots/HERA_FoV_w_strips.png}
  \vspace{-20pt}
\caption{\footnotesize 
HERA will observe a 1440 deg$^2$ stripe centered near $\delta = -30^\circ$. HERA can measure the ionization state around galaxies in, e.g., the GOODS-South field that contains a third of all known $z\!>\!8$ galaxies. HERA's primary imaging data product to the community will be deep cubes along the HERA stripe suitable for cross-correlation.
\label{fig:LSS} }
\vspace{-10pt}
\end{figure}
%  \vspace{-10pt}
%\end{wrapfigure}

\emph{\textbf{Pre-Reionization Heating.}}
\label{sec:EoX} 
Prior to reionization, the 21\,cm signal is a sensitive probe of the first
luminous sources and IGM heating mechanisms. First stars are expected to form
at $z \sim 25-30$ and their imprint on the 21\,cm signal is expected to be
sensitive to the halo mass where they are formed \citep{mesinger_et_al2015}.
The IGM is then expected to be heated by first
generation X--ray binaries
\citep{furlanetto_et_al2006_global,pritchard_et_al2007,mesinger_et_al2013} or
by the hot interstellar medium produced by the first supernovae
\citep{pacucci_et_al2014}, although the heating timing and magnitude is still
very much debated \citep{fialkov_et_al2012}. Dark matter annihilation
\citep{evoli_et_al2014} could also inject energy in the IGM, leaving its
imprint in the 21\,cm signal.

While the observational and analytical state of the art of pre-reionization 21\,cm science is not as advanced as for the EoR \citep{ewall-wice_et_al2016-EoXLimits}, 
with the development of feeds sensitive down to 50\,MHz, HERA will explore the IGM prior to
reionization with a goal to constrain the sources of heating, obtaining, in the case of
X--ray heating, percent-level constraints on the efficiency with which star-forming baryons
produce X-rays \citep{ewall-wice_et_al2015}. Lower
frequency observations also test feedback
mechanisms that interact with low-mass halos at high redshifts
\citep{Iliev_et_al2007,Iliev_et_al2012,ahn_et_al2012}.
Such constraints, while interesting in their own right, also reduce
the susceptibility of the aforementioned $21\,\textrm{cm}$-derived $\tau$
constraints to uncertainties in high-redshift physics \citep{liu_et_al2015}, especially if they are 
combined with upcoming or proposed measurements of 
the pre-reionization sky-averaged spectrum \citep{fialkov_and_loeb2016}. Additionally, they may be crucial to a correct
interpretation of kinetic Sunyaev-Zel'dovich effect constraints on reionization
from the CMB \citep{park_et_al2013}. The high redshift probe of
structure afforded by the low frequency $21\,\textrm{cm}$ measurements will
permit some of the most direct observations of hypothesized suppressions of
small-scale structure
\citep{dalal_et_al2010,tseliakhovich_et_al2011,fialkov_et_al2012} arising from
predicted supersonic relative velocities between dark matter and baryonic gas
\citep{tseliakhovich_and_hirata2010}. No other electromagnetic probe can provide direct observations of this
epoch.

\subsection{Other Scientific Objectives}
\label{subsec:broader_science}

Given its high sensitivity, frequency support, filled aperture and outriggers, HERA is capable of delivering more than the core 21\,cm science discussed above.
HERA will make the data available to the community after only a short quality assurance time period to help facilitate low-frequency science.
Additionally, we encourage groups to leverage HERA as a platform upon which other equipment could be deployed. 
The Moore Foundation's support of HERA feed development and data analysis targeting the pre-reionization science described in \S\ref{sec:EoX}
is a prime example of how we envision this functioning.  Transients also provide an avenue for promising research.
Below, we list examples of the broader science that HERA could impact.

\emph{\textbf{Cross-Correlations with Other Reionization Probes.}}
\label{sec:crossCorr}
HERA's public data provide new opportunities for cross-correlation studies.
Cross-correlation between HERA 21\,cm images and other high-redshift probes
(e.g. JWST; WFIRST; CMB maps; CO, CII, and Ly-$\alpha$ intensity mapping) 
can provide an independent confirmation of the 21\,cm power spectrum
\citep[i.e.][]{lidz_et_al2009,dore_et_al2014,silva_et_al2015,vrbanec2016} and enable rich new studies of
the interaction between galaxies and their ionization environment. 
In particular, cross-correlating 21\,cm with galaxy surveys can measure the characteristic bubble size around galaxies of
different luminosities \citep{lidz_et_al2009} and help separate the degeneracy
between the fraction of photons escaping the galaxies and the
total number of ionizing photons produced \citep{zackrisson2013}.

Fortuitiously, the GOODS-South field---one of the most panchromatically studied regions of the sky, the site of the Hubble UDF, and home to over a third of all known $z>9$
galaxies---lies in HERA's field of view. HERA's images of the IGM can provide environmental context to galaxy surveys through identification of
ionized bubbles ({\em e.g.} \citealt{malloy_lidz2013}), or in a more statistical sense as
described in \cite{beardsley_et_al2015}.

\textbf{\emph{Searching for Exoplanetary Radio Bursts.}}
\label{sec:exoplanets}
HERA could be a powerful tool in the search for bright, highly polarized auroral bursts
from exoplanets %due to the electron cyclotron maser instability 
\citep{treumann2006,hallinan_et_al2015}. These
distinctive, coherent bursts occur periodically at the planetary
rotation rate, which cannot otherwise be measured. 
A burst from a terrestrial planet could point toward
habitability, since it implies a powerful magnetic field
protecting the atmosphere and perhaps the biosphere from energetic stellar wind
particles \citep{tarter_et_al2007}. HERA's sensitivity, large FoV, long campaigns, and precise calibration are all well-suited
for discovering exoplanetary radio bursts from Jupiter-like planets out to 25~pc. Many well studied exoplanetary systems within that distance are in the HERA stripe, including Fomalhaut, Gl 667 C, Gl 433, HD 147513, and Gl 317.

\textbf{\emph{Fast Radio Burst Followup.}}
\label{sec:FRBs}
Fast Radio Bursts (FRB) are millisecond-long radio flashes whose origin has remained a great enigma ever since their discovery \citep{2007Sci...318..777L}. 
HERA could be triggered by nearby, higher-frequency telescopes for FRB followup, saving baseband data and thus full sensitivity to all dispersion measures.
This would require additional hardware, but the architecture allows for such additions.
Bursts like that discovered by \citet{masui_et_al2015}, the lowest frequency
FRB detection to date, should be seen hourly by HERA at 5--10$\sigma$. %(including the effects of scattering)
 Observations at HERA frequencies are very sensitive to the physics of
the intervening medium, particularly deviations from $\lambda^2$
dispersion. Detecting deviations would rule out broad classes of models and could indicate whether FRBs are at cosmological distances.

\section{Measuring the EOR}
\label{sec:eormeas}
Due to the expansion of the universe over cosmological time, we can identify and measure the early Universe via the red-shift of spectral lines.  The hydrogen hyperfine transition at a rest frequency of 1420 MHz is a key spectral line due to the ubiquity of hydrogen and, being a ``forbidden'' transition, the optical depth lets us see through the entire universe back almost to the period of recombination.  The bandwidth
therefore equates to a cosmic distance along the line-of-site of the telescope, with the frequency determining the cosmic age.  An observation of an area of sky over a given bandwidth is therefore providing an average
over a cosmic volume.

As the distance to the EOR is great (about 13 billion light-years) the signal is weak.  However, the signal also subtends the entire sky, so initial measurements of the EOR strive to measure a statistical power spectrum of the signal over the sky since the nature of the reionization process should have a specific spatial signature.  The goal is therefore to measure a range of aggregate spatial scales on the sky, rather than to image the signal directly.  Imaging does remain an ultimate goal to fully understand the process, however we will likely need a greater understanding of the signal characteristics and the systematics of our telescopes to achieve this more difficult goal.

Obviously between our present observation point and the Epoch of Reionization lies the entire intervening universe, which has an intrinsically much brighter signal (up to 6 orders of magnitude or more).  Primarily, this power is due to diffuse Galactic synchrotron radiation, supernova remnants and extragalactic radio sources.  As a first step, areas of the sky where these signals are minimal (for example, outside the galactic plane and away from strong point sources) are targeted \citep{bernardi_2013}.  More importantly and most fortuitously however is the fact that all of these foreground signals are smooth spectrum sources whereas the expected spectrum of the EOR is expected to be rough since it is made up of non-ionized regions which are randomly distributed over a wide range of redshifts.  This fact allows us to try and isolate foregrounds from the EOR, as will be discussed.

The past three years have seen deep EoR observations with PAPER, the MWA, 
the Giant Metrewave Radio Telescope (GMRT; \citealt{paciga_et_al2013}), and LOFAR.
PAPER and the MWA have produced progressively deeper limits
\citep{dillon_et_al2014,dillon_et_al2015,parsons_etal2014,ali_et_al2015}, with PAPER
yielding the first meaningful constraints on the 21\,cm spin temperature during reionization
(Fig. \ref{fig:limits}). %XXX: This figure doesn't really say much about spin temperature.
The inherent challenge is to simultaneously meet stringent sensitivity requirements 
while suppressing foregrounds $\sim5$ orders of magnitude
brighter than the 21\,cm signal, which has been addressed along a number of approaches.  
To address this challenge,  all aspects of the experimental process must be refined and improved, spanning 
calibration ({\em e.g.} \cite{zheng_et_al2014, jacobs_et_al2016, barry_et_al2016}), handling foreground contamination
({\em e.g.} \cite{moore_et_al2013,thyagarajan_et_al2015a,pober_et_al2016}),
and the interferometer design itself ({\em e.g.} \cite{parsons_et_al2012a,dillon_parsons2016}).

A promising approach to make the first detection of the EoR power spectrum 
is a foreground mitigation strategy based largely
on identifying and filtering out a region of phase space where the strong foregrounds are largely confined (Sec. \ref{sec:delayapproach}).  This is the approach
pioneered by PAPER.
By reducing the need
for foreground modeling and subtraction, this approach allowed PAPER to switch to a grid-based antenna layout that
enhanced sensitivity toward fewer Fourier modes and facilitated calibration
based on the ``redundancy" of different antenna pairs measuring the same sky modes \citep{liu_et_al2010,zheng_et_al2014}.
Combined, redundancy and delay filtering provide a robust, inexpensive, and demonstrably successful solution 
to the foreground problem.

Yet PAPER's lack of imaging support and its uneven $uv$ sampling 
leave it with limited diagnostic capability, particularly for direction-dependent systematics
such as polarization leakage from Faraday-rotated emission \Mycitep{moore_et_al2013, nunhokee_etal_2016}.
While concern over such effects has decreased markedly since discovering that
intrinsic point source polarization is lower than previously thought
\citep{asad_et_al2015} and that variable Faraday rotation through the
ionosphere averages down the polarized signal over long observing seasons
\Mycitep{moore_et_al2016,aguirre_etal_2016}, direction-dependent beam effects remain an area of interest.
The MWA's image-based calibration and foreground subtraction strategy provides complementary
capabilities (Sec. \ref{sec:mapapproach}). Imaging with subtraction, while still under development as a viable foreground strategy,
can increase sensitivity by recovering more modes of the cosmological signal (see
Table~\ref{tab:signif}) and help address systematic errors rooted
in the image domain.  HERA's antenna configuration---shown in Figure \ref{fig:arrayConfig} and discussed in \S\ref{sec:arrayConfig}---emphasizes
the proven approaches of redundant calibration and delay filtering, while simultaneously
increasing the extent and density of $uv$ sampling for high-fidelity imaging.

\begin{figure}[t]
	\centering
	\includegraphics[width=.8\textwidth]{plots/summary_results.png}
	\caption{The current best published $2\sigma$ upper limits on the 21cm power spectrum, $\Delta^2(k)$, compared to a 21cmFAST-generated model at $k=0.2$\,$h$\,Mpc$^{-1}$.   Analysis is still underway on PAPER and MWA observations that approach their projected full sensitivities;  HERA can deliver sub-$\text{mK}^2$ sensitivities.}
	\vspace{-10pt}
	\label{fig:limits}
	\label{fig:IGMtemperatureConstraints}
\end{figure}

Despite progress, the fact remains that the 21\,cm EoR signal is intrinsically very faint; making a detection requires a
large instrumental collecting area and a long, dedicated observing campaign.
Although published PAPER and MWA results (Fig. \ref{fig:limits}) do not yet include observations 
at full sensitivity that are still being analyzed (e.g. Fig. \ref{fig:Sensitivities}), 
it is clear already that these instruments lack the sensitivity
to make a conclusive detection (see Table~\ref{tab:signif}). 
HERA addresses this shortcoming with a dish element that delivers a much larger collecting area 
while retaining the necessary characteristics for both proven and forward-looking foreground removal strategies.

This section will provide a brief overview of  the theoretical underpinning of how foregrounds contaminate the measurement (the so-called ``wedge''), discuss the various techniques used to make the measurement and finally provide a brief discussion on calibration issues and techniques.

\subsection{The ``Wedge''}
\label{sec:wedge}
Perhaps the most important advance informing HERA's design is a
refined understanding of how smooth-spectrum foregrounds interact
with instrument chromaticity to produce a characteristic ``wedge" of
foreground leakage in Fourier space (see Fig.~\ref{fig:wedge}), 
outside of which the 21\,cm signal dominates (the ``EOR window'').  
Through theoretical and observational work
\citep{datta_etal2010,morales_et_al2012,parsons_et_al2012b,vedantham_2012,thyagarajan_et_al2013,hazelton_et_al2013,pober_etal2013b,liu_et_al2014a,liu_et_al2014b},
we have learned how the boundary between the wedge and  the EoR window is determined by the separation between antennas,
signal reflections within antennas, and the angular response of the antenna beam.  Deep integrations also show us
that, to the limits of current sensitivity, foreground emission is absent outside of the wedge; it can only 
appear there through instrumental leakage \citep{parsons_etal2014,ali_et_al2015,moore_et_al2016,kohn_et_al2016}.

The power spectrum measurement provides the spatial correlations across the sky, characterized by the magnitude of the wavenumber, ${\bf k}$.
Though the full magnitude of the {\bf k}-vector is used, it is instructive to split it into two components,  $\kvec =  \kvpr + \kpar{\hat {\bf z}}$ where $|\kvpr|\equiv \kpr = 2\pi b/\lambda X$ is determined by the antenna baseline ($b$) and $\kpar=2\pi/(YB)$ by the bandwidth ($B$).  Here $X$ and $Y$ are cosmological parameters relating angular size and spectral frequency to cosmic volumes respectively (so, relating wavenumber to physical volume at a given redshift).  This is useful since it allows us to split out the chromatic response of the interferometer visibility measurement and isolate a phase space where foreground sources ({\em i.e.} everything {\em not} the EOR) contaminate the signal of interest from where they don't.

The $\kvpr$ components are directly proportional to the baselines and $\kpar$ are proportional to the Fourier transform of the frequency response.   The Fourier transform of a frequency spectrum is a delay, hence the name ``delay-spectrum'' for the technique that uses this approach.  Note that cosmic evolution limits the largest bandwidth (which determines the smallest $\kpar$) to about 10 MHz -- for larger bandwidths the evolution of the Universe begins to impact the result.  For HERA, wavenumbers are dominated by the bandwidth, not the baseline.

Figure \ref{fig:kperf} shows many of the dependencies on wavenumber by the bandwidth, configuration and cosmos by plotting the perpendicular wavenumber ($\kvpr$; black lines), parallel wavenumber ($\kpar$; blue lines) and total wavenumber ($k$; green lines) as a function of redshift for various bandwidths and baselines.  The redshift range is appropriate for the extended frequency range that is the goal for HERA, however $z$=6-13 is the primary region for EOR studies.
The lines are labelled for the assumed bandwidth and/or baseline, depending on the type of wavenumber shown.  Obviously many values may be used, but here a small sampling of values appropriate for HERA are shown.    
Wavenumbers with an assumed bandwidth ($k$ and $\kpar$ are shown as stepped profiles, with the redshift steps equal to the bandwidth at that redshift.
Note that for 100 MHz, the concept of a power spectrum at a given redshift is no longer appropriate, which is why the line is dashed.
Also in Fig. \ref{fig:kperf}, the lime-green shaded area indicates the initial region in which HERA intends to detect and characterize the EOR using the delay-spectrum technique.

\begin{figure}[h!]
\centerline{
\includegraphics[width=.4\textwidth]{plots/kperf.pdf} 
}
\caption{\small $|\kvec |, \kpar, \kperp$ wavenumber plots at various bandwidths ($k_{||}$; blue), baselines ($\kperp$; black), and total ($|\kvec |$, green) as a function of redshift.  The stepped profiles for $k_{||}$ and $|\kvec |$ are the line-of-site redshift ranges for that bandwidth (too small to see at 1 MHz).  See text for discussion.  The full redshift range is the HERA extended goal.  The initial HERA delay-spectrum power spectrum goal is the light green square region in the upper left.}
\label{fig:kperf}
\end{figure}

In these variables, the ``contaminated'' phase space is a wedge-shaped region in $\kpr$:$\kpar$ space such that the ratio of the time-delay across a given baseline ($b/c$,  which impacts the chromatic visibility) and the delay associated with a given bandwidth ($1/B$) is less than some parameter determined by the details of the system, which we denote as $1/\beta$.  Substituting $\kpr$ and $\kpar$ in for $b$ and $\Delta\nu$ in this ratio, this wedge is bounded by
\begin{equation}
\kpar \le \beta\frac{X\lambda}{Yc}\kpr + \frac{S}{Y}
\end{equation}
where an offset parameterized by $S$ accounting for affects related to the combined spectral smoothness of the foregrounds and the antenna response has been included(see Fig. \ref{fig:wedge}).

If this $\beta$ factor is determined by the chromaticity of the longest baseline, we see that $\beta=1$ and this line is referred to as the ``horizon line'' ({\em i.e.}, the delay of a source at the horizon).  Systematics will push that line such as to close the EOR ``window'' to the upper left ($\beta > 1$).  If we can completely control systematics and constrain the affects of the foreground to {\em e.g.} the main beam, the line will move to decrease the size of the wedge ($\beta<1$).
Therefore, a key issue to measuring the EOR power spectrum is to understand and minimize systematics.  For further discussion and an analytical derivation, see \cite{zahn_etal2012,vedantham_2012,liu_et_al2014b},  in simulation see \cite{datta_etal2010,hazelton_et_al2013} and for observations
\cite{pober_etal2013b,2015arXiv150601026P,parsons_etal2014,2015arXiv150206016A}.

Note that the unit of the spatial wavenumber $k$ is 1/length, in this case the relevant length scale is megaparsecs (Mpc; 1 Mpc = 3.086$\times10^{22}$ m).  In order to normalize for updated measurements of the Hubble constant, it is further normalized by a factor $h$, where Hubble's constant is expressed as $H_0=100h$ (km/s)/Mpc , such that the wavenumbers are expressed as $h$/Mpc.  Also note that at the red-shifts of interest $X$ has a value of about 160 Mpc/deg and $Y$ has a value of about 16 Mpc/MHz.

\begin{figure}[t]
\centerline{
\includegraphics[height=3in]{plots/wedges.pdf}
}
\caption{The EoR ``window'' and foreground ``wedge'' are shown in this plot, which splits the wavenumber into $\kvpr$ (horizontal) and $\kpar$ (vertical) components.
Smooth-spectrum foregrounds with an intrinsic spectrum indicated by the gray bar along the bottom create the ``wedge'' when they interact with the interferometer's chromatic response, shown with a foreground image from the MWA.  Systematics can extend the wedge to the upper left.  The EoR window (shown with a simulation
from \citealt{mesinger_et_al2011}) to the upper left should be free from foregrounds.}
\label{fig:wedge}
\end{figure}

The techniques to measure the power spectrum may be broken down into two principle techniques -- delay-space (\S\ref{sec:delayapproach}) and map-making (\S\ref{sec:mapapproach}) -- plus additional hydrid methods.  These are briefly summarized below, with an emphasis on the delay-spectrum technique as the initial approach taken by HERA.  Finally, we provide a short discussion on calibration in the concept of this design.

\subsection{Delay-Spectrum Approach}
\label{sec:delayapproach}
Given that the response of an interferometer natively measures the power in Fourier modes of the sky within its beam, we see that it is a natural instrument to use for this measurement of the EOR spatial power spectrum.  The delay-spectrum approach leverages the interferometer measurement to optimize sensitivity to the desired modes while rejecting modes contaminated by the foreground power in the wedge.  Other than potentially handling overlapping bins in {\em uv} space, the delay-spectrum approach does not combine baselines before squaring and calculating the power spectrum, which contrasts to the map-making techniques briefly described in the following section.

We can essentially relate the sky power spectrum mode ${\hat P}(\kvec)$ to an interferometer baseline visibility ${\tilde V}(\bvec)$, which is the equivalent of the (complex) fringe pattern in a double slit experiment:
\begin{equation}
{\hat P}(\kvec) \approx\frac{X^2Y}{4k_B^2}   \left[\frac{{\tilde V}^2(\bvec)}{\Omega_b B/ \lambda^4} \right]
\label{eq:Pk_baseline}
\end{equation}
where $X$ and $Y$ were introduced above, $\Omega_b$ is the integrated beam response, $B$ is the effective bandwidth, $\lambda$ is the observation wavelength, and $k_B$ is Boltzmann's constant.  The terms in square brackets are instrumental terms, as opposed to the constants and cosmological parameters.  
Rather than ${\hat P}(\kvec)$, the literature generally works with a volume-normalized parameter given by $\Delta^2(k) = \frac{k^3}{2\pi^2}{\hat P}(\kvec)$.

Since the thermal noise per visibility baseline may be expressed as
\begin{equation}
V_N = \left(\frac{2k_B}{\lambda^2}\right)\left(\frac{T_{sys}}{\sqrt{2B\tau}}\right)\Omega_b B
\label{eq:sensitivity_per_baseline}
\end{equation}
where $\tau$ is the integration time (see {\em e.g.} \citealt{thompson_et_al2001}),
to determine the sensitivity of an instrument to the power spectrum per baseline we can substitute the thermal noise per baseline for the visibility in Eq. \ref{eq:Pk_baseline}.  Further, since an interferometer typically measures many baselines which may be averaged together to improve the signal-to-noise per $k$-bin, the total sensitivity may be approximated as
\begin{equation}
\Delta^2_N (k,z)\approx \frac{k^3X^2(z)Y(z)}{4\pi^2} \left[\frac{T_{sys}^2(z)\Omega_b(z) }{\tau(z) \mathcal{N}_s(k)}\right]
\label{eq:sensitivity}
\end{equation}
where $\mathcal{N}$ represents the improvement in sensitivity based on the array configuration, which may significantly boost the sensitivity if optimized for this measurement (see \citealt{parsons_et_al2012b}).  Dependencies on redshift and wavenumber have been included.

Fig. \ref{fig:boost} (left) plots this ``redundancy boost'' factor for various configuration types relative to an array in a 19-element hexagonal configuration.  They are labelled for the type of configuration (hex, grid, imaging) and the number of antennas (from 37 to 350).  Note that (a) {\em hex} is based on the configuration used for HERA (14.6m center-to-center spacing); (b) {\em grid} for that of PAPER (4m N-S, 16m E-W spacing); (c) {\em imaging-128} for MWA \citep{tingay_et_al2013}; and (d) {\em imaging-48} for LOFAR (\citealt{2013A&A...556A...2V}; note that this assumes the ability to fully correlate the HBA-pairs).  The plot shows that for a fixed number of elements using the delay-spectrum approach, redundant arrays (hex and grid) provide about an order of magnitude improvement over imaging arrays.  Since the baselines go as order $N^2$, this corresponds to using about 1/3 of the number of elements to yield the same performance (as verified by the Hex-37/Imaging-128 lines).  This advantage does not extend to the other approaches discussed below and imaging configurations would be preferred.  Note that a filled hexagonal-packed array provides excellent imaging as well, but with potentially limited resolution unless outriggers are included.

Another approach to improving sensitivity is to make the collecting area per element larger.  In the limit, this yields a fully filled aperture.  A larger element does limit the accessible field-of-view, but this is not a huge liability for a focussed experiment as long as an appropriate patch of ``cold'' sky passes within its field-of-view.   Fig. \ref{fig:boost} (right) shows the relative sensitivity incorporating configuration and element size of various arrays relative to HERA-19, calculated for a redshift corresponding to the peak signal at a hydrogen reionization fraction of 0.5.  The circles to the right indicate the relative sizes of the element.  Note that MWA and LOFAR are cross-hatched to indicate phased array tiles, which uses small antennas to phase up a full element response.

\begin{figure}[h!]
\centerline{
\includegraphics[height=6cm]{plots/boost.pdf}
\includegraphics[height=6cm]{plots/sensratio.pdf}
\includegraphics[height=4cm]{plots/elementSizes.pdf}
}
\caption{Left: Representative boost factors ($\mathcal{N}$ in Eq. \ref{eq:sensitivity}) normalized to a 19-element hexagonal configuration spaced at 14.6m.  For a given element count, one can get about an order of
magnitude improvement in delay-spectrum sensitivity.  Alternatively, for the same sensitivity one can build roughly 1/3 the antennas.  Right:  Power spectrum sensitivities for various arrays normalized to HERA-19.  Here the element size is included, which is shown as an appropriately sized circle.}
\label{fig:boost}
\end{figure}

As an aside, note the parallels to imaging with an interferometer.  For imaging, one takes a 2-D spatial Fourier transform of the visibilities, and produces images as a function of frequency.  Here, one is taking a 1-D Fourier transform along the other axis (the bandpass) and produces a 3-D spatial power spectrum.  In imaging, point sources are transformed to a single pixel, whereas for the delay-spectrum, smooth sources are transformed to a single channel (but then get forced into the wedge due to the inherent chromatic response of an interferometer).  For imaging, we are concerned with ripples in the passband due to standing waves, whereas for the delay-spectrum we are concerned with delayed versions of the signal due to multiple reflections.

\subsection{Mapmaking Approach}
\label{sec:mapapproach}
In contrast to the delay-spectrum approach, mapmaking approaches combine baselines to build up information before squaring and calculating the power spectrum.  The conceptually simplest approach is to first make an image cube of the sky as a function of angular position and frequency, and then to Fourier transform, square, and bin the cube to determine the power spectrum \citep{liu_tegmark2011,dillon_et_al2013a}. The image domain is a natural place to combine information from partially-coherent pairs of visibility measurements (either from different baselines or from the same baseline at different times). Mapmaking can be a lossless data compression step \citep{tegmark1997b}, though keeping track of statistical properties of the maps, which have complex frequency and position dependent point spread functions and noise covariance matrices, can be computationally challenging \citep{dillon_et_al2015a}. Existing approaches have had to make a number of approximations, including that point spread functions do not vary over the field of view and that noise is not correlated between $uv$-cells. Additionally, mapmaking is in principle not as vulnerable to polarization leakage as the delay-spectrum approach, since only polarization mismodeling (rather than any kind of polarization asymmetry) can cause leakage from Stokes $Q$ or $U$ to $I$.

%This approach is being pursued primarily via FHD.
%XXX: this needs the be expanded.

One advantage of such an approach is that sky images can be interesting in their own right, both for dealing with measurement systematics and for accessing the non-Gaussian observational signatures (such as ionized bubble structures) that are missed by variance statistics such as the power spectrum. Mapmaking also allows for the direct subtraction of bright foregrounds, potentially expanding the EoR window. However, accessing these high-order statistics or the modes of the power spectrum inside the wedge using a mapmaking approach requires extreme (and so-far undemonstrated) precision in calibration and forward-modeling of bright foreground models through instrument systematics.

\subsection{Hybrid Approaches}
\label{sec:hybridapproach}

While they were presented separately above, the delay-spectrum and map-making approaches need not be viewed as mutually exclusive. For example, mapmaking need not be limited to the real space image basis; pipelines that combine multiple baselines into estimates of various Fourier amplitudes of the sky are formally also mapmaking pipelines. Like the delay-spectrum approach, such pipelines (see, e.g., \citealt{trott_et_al2016}) avoid the image domain entirely. This helps to prevent artifacts that may be introduced by imaging algorithms.

More generally, it is possible to express a mapmaking algorithm as a linear operator that acts on visibility data to produce a compressed dataset \citep{dillon_et_al2015a}. It follows then that squaring this to obtain an estimate of the power spectrum then results in a weighted quadratic combination of visibilities, where one multiplies the visibility from every baseline with that from every other baseline. These visibility product pairs are subsequently normalized to form their own individual estimates of the power spectrum before all the normalized pairs are summed together to form a final power spectrum. Measuring a power spectrum in this way combines the best aspects of the mapmaking and delay-spectrum approaches. Power spectra are formed directly by the cross-multiplication of visibility data, preserving the delay-spectrum approach's strategy of staying close to quantities measured by an interferometer; on the other hand, by multiplying together the data from every possible combination of baselines, one retains all the information from partially-coherent visibility pairs captured by mapmaking. Indeed, it can be shown \citep{liu_et_al2014a} that in the limit of infinitely fine Fourier space bins, the hybrid method that we have just described is mathematically equivalent to the mapmaking approach. In practice, the hybrid method comes with the added benefit of allowing particular baseline pairs that are suspected of being affected by instrumental systematics to be more surgically downweighted.

Fundamentally, the hybrid approach and the optimal mapmaking approach \citep{dillon_et_al2015a} are very similar. Whether it makes computational sense to examine all pairs of partially-coherent visibilities or to perform mapmaking as an intermediate data-compression step before power spectrum estimation (despite the complexity of the resultant map statistics) is an open question that depends on a number of factors, including the number of elements in the array, its layout, and its integration time.
However, similar to map-making techniques, these approaches need additional precision in calibration and modeling development in order to be implemented with confidence.

\subsection{Calibration}
\label{sec:calibration}
As examined above, maximizing sensitivity to a limited number of spatial modes for the delay-spectrum technique requires a grid of antennas that simultaneously measures the same baselines using many redundant pairs of antennas. This redundancy may also be exploited to calibrate the array as well, as long as the true antenna positions fall close enough to that ideal grid \citep{liu_et_al2010}. This technique was pioneered with the MITEoR experiment \citep{zheng_et_al2014}, and has resulted in a package called OMNICAL developed for PAPER and HERA.  \citet{ali_et_al2015} provides a description of this technique as it applies to calibrating the PAPER array, which is mostly situated on a grid with many redundant baselines.  This section provides a high-level overview of calibration as it applies to HERA.

Fundamentally, the problem of calibration is the determination of a complex, frequency-dependent gain for each antenna which arises due to differences in amplifiers, cables, etc. Given those gains, the visibility measured by antennas $i$ and $j$ at frequency $\nu$ is given by 
\begin{equation}
V_{ij}^\text{measured}(\nu) = g_i(\nu) g_j^*(\nu) V_{ij}^\text{true}(\nu), 
\end{equation}
where each $V_{ij}(\nu)$ is the measured or true visibility and $g_i(\nu)$ is the gain on antenna $i$ at frequency $\nu$.\footnote{This discussion ignores electromagnetic cross-talk between antennas and antenna-to-antenna variation of the primary beam.} With its highly-redundant array configuration (see \S\ref{sec:arrayConfig}), HERA's 350 antennas measure 61,075 visibilities, but only 6,140 unique baselines. Given a set of measured visibilities, it is possible estimate both the 6,140 unique true visibilities and the 350 complex gains simultaneously because that system of equations is greatly overdetermined. A method for linearizing the system and applying standard minimum-variance linear estimators was developed by \citet{liu_et_al2010} and first applied in \citet{zheng_et_al2014}. Unlike traditional calibration of radio interferometers, redundant calibration makes no reference to any sky model, requiring merely that the sky is bright relative to the noise on each antenna (which is always true in the sky-noise dominated regime characteristic of 21\,cm observatories with wide fields of view).

However, at each frequency, degeneracies in the redundant calibration procedure prevent it from solving four numbers per frequency for the whole array (down from 700 in HERA's case if full calibration, as opposed to redundant calibration, were needed). The first is an overall gain---one can multiply each $g_i(\nu)$ by 2 and divide each $V_{ij}^\text{true}$ by 4, for example, and get exactly the same measured visibilities. The others are an overall phase and two other phase terms that describe the tip and tilt of the array. These degeneracies mean that the output of any redundant baseline calibration also need to undergo a final, ``absolute'' calibration. Absolute bandpass calibration has thus far been performed for PAPER and MITEoR using a sky model, though it remains an open question whether this reduced set of calibration parameters can be further constrained without an explicit sky model.

\section{High-Level Requirements} 
\label{sec:requirements}
From the science and techniques discussed above, we may derive some high-level requirements predicated on optimizing for the delay-spectrum approach of detecting the EOR power spectrum while not unduly limiting other approaches or other science goals.  The fact that HERA is an experiment and not meant to be a long-term general use facility greatly facilitates this optimization. This section describes the element, configuration, frequency and sensitivity requirements that largely set the overall design concept.

\subsection{General Antenna Design and Configuration}
The primary new feature of HERA over previous generation experiments (in this context primarily PAPER) is the use of large elements to increase the sensitivity, as indicated in Fig. \ref{fig:boost}.  The cost-performance optimization will 
be discussed below, however we see that to address the central challenges of the delay-spectrum approach in the context of the wedge.
HERA should use close-packed antennas that
minimize signal reflections over long delays and deliver significant forward gain relative to their horizon response.
Tests with prototype HERA antennas (Figs.  \ref{fig:delayspec} and \ref{fig:orbcommexptandbeammap}, discussed in \S\ref{sec:antenna})
indicate that a moderately large parabolic dish with a short focal height can meet these requirements
\citep{ewall-wice_et_al2016-EoXLimits,neben_et_al2016,thyagarajan_et_al2016}.  Note that to pursue map-making and hybrid techniques, additional antenna outriggers are valuable additions, so the design
should be able to accommodate $\approx$1km baselines.  Note that in referring to the EOR sensitivity of the array, often the number of antennas just in the core are
indicated -- that is for EOR HERA-320 and HERA-350 are often used interchangeably.

\subsection{Frequency and Bandwidth}
\label{sec:freqs}
Figure \ref{fig:cosmos} indicates the frequency range requirement to probe the expected timescale of the epoch of reionization using the 21cm line of hydrogen as the probe.  
These limits are derived from to-date complementary probes of reionization which include measurements of
the optical depth to last scattering in the CMB, QSO spectra, Ly-$\alpha$
absorption in the spectra of quasars and gamma ray bursts and the demographics
of Ly-$\alpha$ emitting galaxies
(Figure~\ref{fig:IonHist}). 
Constraints from these probes are still weak:
Ly-$\alpha$ absorption saturates at very small neutral fractions; galaxy
surveys directly constrain only the bright end of the luminosity function and
depend on an unknown escape fraction of ionizing photons to constrain
reionization; CMB measurements probe an integral quantity subject
to large degeneracies. Even when these observations are
combined into a single 95\% confidence region, the bounds remain weak.
For example, $x_{HI}$ spans almost the entire allowable range of [0,1]
at $z=8$. 21\,cm reionization experiments place much tighter constraints on ionization, with the
red band showing the forecasted 95\% confidence region derived from HERA data,
after marginalizing over astrophysical and cosmological parameters.

From these measurements and models, HERA is required to cover a redshift range of $z=$6 to 13, corresponding to a frequency range of 101 - 203 MHz.  We adopt 100-200 MHz for full performance as the requirement.  Extending the science to include the Dark Ages remains a goal so that efforts are on-going to design a feed to increase the lower limit down to at least 70 MHz without compromising the performance in the above requirement range.  The fall-back position for low frequencies is a serial deployment of a scaled version of the HERA feed or to potentially build additional elements specifically for low frequencies.

The scientific requirement on channel bandwidth is to allow access to k-modes of $\sim1$ h/Mpc, which requires about 256 channels over the 100 MHz bandwidth.  However, in order to handle radio frequency interference as well as to allow the bandpass to be characterized, a specification of 1024 channels has been chosen.  This yields a channel bandwidth of 97.7 kHz.

The total simultaneously processed bandwidth is the full 100 MHz, since we wish to efficiently probe the entire redshift range.  The main impact from this requirement is the bandwidth from the digitizer back to the correlator, however this bandwidth is easily handled with current generation digital back-ends.

\subsection{Delay Response}
\label{sec:delayspec}
Early analysis \citep{elementmemo} indicated that to trace the EOR over a wide range of redshift, internal reflections should be attenuated by about 60 dB from the initial incoming wave by about 60 ns, or a VSWR $<$ 1.002 for frequencies $<$ 17 MHz.  This provided the basis for the initial design of the element.    For the contemplated feeds on a primary focus antenna, it was found that a focal length/diameter radio ($f/D$) of approximately 0.32 was optimal.  We may therefore estimate the attenuation as a function of delay and diameter and assign this a ``cost function''.

Figure \ref{fig:costfig} summarizes the adopted cost function.  
The vertical axis shows a very steep delay ``cost'' at 60 ns and the horizontal axis is the cost/performance model, as described in the following section.  
The color coding is the product of the two assigned costs, with red showing a high cost normalized to 350 14m elements (clipped at a ratio of 1.25) and purple being the minimum.
The curves labelled `1', `2' and `3' are the round-trip travel times for 1, 2 and 3 reflections between the feed and the vertex.
To obtain 60 dB by 60 ns we may interpret those lines in one of two ways:  
(a) at a fixed attenuation/reflection, what diameter is required, and
(b) at a fixed diameter how much attenuation per reflection can be allowed.
For 1 round-trip, for the assumed $f/D$ over the diameters of interest, from Fig. \ref{fig:costfig} all delays are less than 60 ns, so this provides no constraint.   If we want to be able to handle 2 round-trips worth of delay, the diameter should be less than 14.2m, which is near the cost minimum.  If we assume 14.2m, then we can estimate that each reflection must average a return loss better than 15 dB, an aggressive target.  And finally, if we need to allow for 3 round-trips, our antenna diameter would have to be less than 9.4m, where the cost is quickly rising.  To handle the delay requirement therefore, the diameter should be less than 14.2m, or the return loss specification becomes very aggressive.

As a key specification, this has been a focus of study with a series of prototypes and analyses based on modeling, and measurements have been conducted to provide much more rigor than this early specification \citep{ewall-wice_et_al2016-EoXLimits,neben_et_al2016,thyagarajan_et_al2016,patra_et_al0216}.  The simple specification set above more appropriately becomes a series of curves for attenuation and delay at different wavenumbers for realistic foreground models.  Fig \ref{fig:delayspec}, adapted from \cite{thyagarajan_et_al2016}, shows the results at $k$ = 0.10, 0.15, and 0.20 $h$/Mpc for the adopted design.  The analysis shows that
the dish design for should should allow for EOR detections for $k>0.10 h/$Mpc.  We also see that the initial simple ``60@60'' specification was a overly conservative, however it served as a good initial goal.

\begin{figure}[h!]
	\centering
    \includegraphics[width=0.49\textwidth]{plots/delayspecplot.pdf}
     %\includegraphics[width=0.5\textwidth]{plots/ps1d_with_delay_kernel.pdf}
	%\caption{Left:  
	\caption{Antenna delay attenuation specification with realistic foregrounds from \cite{thyagarajan_et_al2016} along with the initial conservative ``60@60'' specification.  HFSS calculations (black line) show the HERA dish meeting the specification.  Measurements on the Green Bank prototype are shown in  \citep{patra_et_al2016}.
}
%Right: Modeled foreground power spectra with (green) and without (blue) the antenna chromaticism over a 10\,MHz subband centered at $z=7.5$. The spectral smoothness of the dish enables foreground isolation below the level of a fiducial reionization signal (black) for most $k_{\parallel}$ modes, yielding percent-level constraints on model parameters.}
	\label{fig:delayspec}
	\vspace{-10pt}
\end{figure}

\subsection{Sensitivity Optimization}
\label{sec:cost}
As an instrument to characterize the power spectrum over the evolution of the EOR, the specification on sensitivity is to make at least a nominal detection over the redshift region of support ($z$=6-13) along with a very robust detection at the peak.  As seen in Fig. \ref{fig:Sensitivities} and Table \ref{tab:signif}, an array comprising 320 14-m core antennas (labelled HERA-350, which includes the outriggers) has sensitivity to a
fiducial reionization model across that redshift range.  We therefore adopt the delay-spectrum power spectrum sensitivity of 320 14-m close-packed antennas as the minimum sensitivity needed.  
For this sensitivity, the number of elements needed for a given diameter is
\begin{equation}
\label{eq:ND}
N_{\text core} = (320*14)/D
\end{equation}
where $N_{\text core}$ and $D$ are the number of antennas in the core and their diameter respectively.  Note that the linear number-diameter dependence was derived by running multiple sensitivity codes ({\em e.g.} 21cmSense) over a range of values for the HERA configuration for wavenumbers greater than 0.15.  This dependency differs from \cite{2013ExA....36..235M} and analytical expressions, mostly likely due to the assumed configuration.
Optimization is 
therefore done for this sensitivity against cost for various diameters and total element numbers, while trying to minimize multiple reflections on long delays.

The cost function in Fig. \ref{fig:costfig} is shown as the colored background and color bar, with the delay-spec vertically and the cost at a fixed sensitivity horizontally. 
To determine the optimal diameter, full system costings were done on a range of diameter sizes from 5m - 25m, with the total number set by Eq. \ref{eq:ND}.  This included the full system, except the post-processing (that is, everything after the archived correlator output).  
The resulting normalized system cost/performance-curve  becomes very flat (colored dark blue/purple) for diameters greater between about 12 and15 m, which is consistent with the chosen value of 14 meters based on other system considerations.

\begin{figure}[t]
\centerline{
\includegraphics[width=0.6\textwidth]{plots/costfig.pdf} 
}
\caption{\small The background coloring with colorbar shows the relative cost/performance for delay and cost at a fixed performance as a function of element diameter.  See text and \cite{elementmemo} for details.  A diameter of 14-m (vertical dashed line) is near the cost minimum and consistent with assumption for the delay specification.
\label{fig:costfig}}
\end{figure}

\section{System Design} \label{sec:design}
As described in the previous section, the critical insights from the first generation 21\,cm EoR experiments have been applied to define
the requirements for HERA---an instrument designed to ensure that foregrounds remain bounded
within the wedge while delivering the sensitivity for
high-significance detections of the 21\,cm reionization power spectrum with
established foreground filtering techniques
\citep{pober_et_al2014,greig_and_mesinger2015}.
In this section, we summarize key features of the HERA design (see Table~\ref{tab:BasicParameters}) 
and system architecture (see Fig.~\ref{fig:overallBlockDiagram}).
This architecture directly inherits from the PAPER and MWA experiments; HERA begins by reusing the
analog, digital, and real-time processing systems deployed for PAPER-128.  As HERA develops, this
architecture is incrementally upgraded to improve performance and add features while simultaneously
addressing issues of modularity and scalability.  As with PAPER, HERA proceeds in stages of development,
with annual observing campaigns driving a cycle of development, testing, system integration, calibration, and analysis.
This cycle ensures that HERA's instrument is always growing, that systematics are being found and eliminated at the
earliest build-out stages, that data analysis pipelines are tested and debugged while data volumes are smaller,
and that HERA is always producing high quality science.

Note that in addition to the 19 elements in South Africa (and an initial prototype near Berkeley, CA), a pair of elements at the Green Bank Observatory in West Virginia have been used extensively in the validation testing, discussed below.  Also, three elements are currently under construction at the Mullard Radio Astronomy Observatory outside Cambridge, UK for additional feed and element tests.

For details of the first generation HERA system (the existing and commissioned PAPER signal path), see \cite{parsons_et_al2010}.  This section will discuss the new system to be deployed beginning and 2016 and tested in conjunction with the previous system.  Fig \ref{fig:overallBlockDiagram} provides a system block diagram of the new HERA architecture.

\begin{table}[t]
\small
\Caption{0.0in}{0.99}{-0.1in}{HERA-350 design parameters and their observational consequences. Angular scales computed at 150\,MHz.}
\label{tab:BasicParameters}
\begin{center}
\begin{tabular}{l | l}
\multicolumn{1}{c}{\emph{\textbf{Instrument Design Specification}}} & \multicolumn{1}{c}{\emph{\textbf{Observational Performance}}}\\
\hline
\textbf{Element Diameter:} 14\,m & \textbf{Field of View:} 9\arcdeg \\
\textbf{Minimium Baseline:} 14.6\,m & \textbf{Largest Scale:} 7.8\arcdeg\\
\textbf{Maximum Core Baseline:} 292\,m & \textbf{Core Synthesized Beam:} 25\arcmin\\
\textbf{Maximum Outrigger Baseline:} 876\,m & \textbf{Outrigger Synthesized Beam:} 11\arcmin\\
\textbf{Frequency Range:} 50--250\,MHz & \textbf{Redshift Range: $4.7 < z < 27.4$} \\
\textbf{Frequency Resolution:} 97.8\,kHz & \textbf{LoS Comoving Resolution:} $1.7$\,Mpc (at $z=8.5$)\\
\textbf{Survey Area:} $\sim 1440$ deg$^2$ & \textbf{Comoving Survey Volume:} $\sim 150$\,Gpc$^3$ \\
$\mathbf{T_\textbf{sys}}$: $100 + 120 (\nu/\rm{150~MHz})^{-2.55}$ K & \textbf{Sensitivity after 100\,hrs:} 50 $\mu \rm{Jy}~\rm{beam}^{-1}$  \\
\hline
\end{tabular}
\end{center}
\end{table}

\begin{figure}[h]
	\centering
	\includegraphics[width=1\textwidth]{plots/HERA_high_level_block_diagram.pdf}
	\caption{HERA's signal path.  Front-end amplifiers at the antenna feed drive signals on short coaxial cables to 
field nodes.  Nodes contain post-amplifiers and Smart Network ADC Processor (SNAP) boards that digitize, channelize,
and packetize data for optical transmission in 10 Gb Ethernet format.  Optical fibers are aggregated in a field container
onto a 10 km fiber bundle connecting to the Karoo Array Processing Building, where signals are cross-multiplied
in the X processor.  After correlation, visibilities are stored by the Librarian, compressed and redundantly calibrated
by the Real-Time Processor, and transmitted over the network to clusters for storage and analysis.  Final products are
hosted on public-facing NRAO servers, with a web interface for selecting and downloading data.}
	\label{fig:overallBlockDiagram}
	\vspace{-5pt}
\end{figure}

%\subsection{Site}
%\label{sec:site}
HERA is located at the Karoo Astronomy Reserve.

\subsection{Antenna}
\label{sec:antenna}

As discussed in the previous section, the design of the design principles are three-fold:
\begin{itemize}
\item optimize for the delay-spectrum technique of measuring the EOR power spectrum,
\item minimize costs, and
\item the experiment has a limited lifetime of about five years.
\end{itemize}
The first item primarily means that chromatic effects corresponding to delays appropriate for the measurement described above must be below the expected signal level, which essentially determines the focal length.  The second item constrains the diameter and element count, as well as the focal length over diameter ratio ($f/D$), based on a cost function and maximizing sensitivity per element.  And the third item constrains the construction materials and methods and the operational model.  This led to a fixed transit element with a diameter of 14 meters to strike an optimal balance between sensitivity and systematics.

The large collecting area of a HERA element yields nearly 5 times the sensitivity of an MWA tile and more
than 20 times that of a PAPER element, but it does so without substantially
degrading our ability to isolate and remove foreground emission on the basis of
spectral smoothness (Fig. \ref{fig:reflectometry}, right panel).  As shown in
\Mycitet{parsons_et_al2012b}, the amplitude and timescale of signal reflections
relates directly to the leakage of smooth-spectrum foregrounds into regions of Fourier space 
used to measure reionization, as discuss in Sec. \ref{sec:wedge}.
To facilitate foreground filtering, HERA's antenna element is designed to suppress reflections at long time delays.

As shown in Fig.~\ref{fig:delayspec}, HERA's element meets this requirement in
reflectometry tests, performing comparably to the PAPER element.  The
relatively short (4.5\,m) focal height of the parabolic dish 
minimizes the timescale of feed-dish reflections while maintaining the feed's
illumination of the dish.  HERA's feed design is based on the PAPER sleeved
dipole element incorporating a cylindrical backplane to
improve symmetry in the polarized beam response patterns (Sec. \ref{sec:feed}, which also presents HFSS calculations of the full feed/element system).

Figure \ref{fig:orbcommexptandbeammap} shows the beam pattern measured at 137 MHz with a beam mapping
system using the ORBCOMM satellite network \Mycitep{neben_et_al2016}.  Results
indicate an effective per-element collecting area of 93\,m$^2$, compared with the
theoretical maximum Airy response of 155 m$^2$.  The measured primary beam is
consistent with simulations at the 0.1\% to 0.5\% level, with a full width at
half maximum of $\sim$$10^\circ$, and a first sidelobe at -20 dB \citep{ewallwice_et_al2016,neben_et_al2016,patra_et_al2016,thyagarajan_et_al2016}.
Feed-to-feed coupling between adjacent antennas has been measured to be below -50 dB, indicating
that mutual coupling will not be problematic.

\begin{figure}[h!]
	\begin{tabular}{ll}
	\begin{minipage}{4.2in}
\includegraphics[width=2.1in]{plots/ref_dipole_and_hera_dish.jpg}
\includegraphics[width=2.1in]{plots/orbcomm_dish_beam_map_530cm_feed.pdf}
	\end{minipage} & 
	\begin{minipage}{2.05in}
	\caption{Left: The first of three prototype dishes at NRAO--Green Bank, used for measuring beam frequency structure with reflectometry and the beam pattern at 137\,MHz by comparing satellite signals to the reference dipole in the foreground \citep{neben_et_al2016}. Right: The measured EW power pattern plotted with dashed lines marking zenith angles of 20$^\circ$, 40$^\circ$, 60$^\circ$, 80$^\circ$.} 
	\label{fig:orbcommexptandbeammap}
	\end{minipage}
	\end{tabular}
	\vspace{-15pt}
\end{figure}

\subsubsection{Element Construction}
To minimize cost and aid in construction in a remote location, the construction materials are readily available standard construction materials, making for a very cost-effective design.  The feed is supported from three utility poles using line and hardware primarily from boating activities which can handle the load and external environment.  Given the close-packed design, each pole (except for the perimeter) are shared between three antennas.  This both reduces the number of poles, but also makes for a balanced load.  Note that the perimeter poles can be stayed with guy lines if needed.   The rim is also supported off of these poles, with three smaller posts in between such that the outer perimeter is regular dodecagon shape.  Two of the three intermediate posts are shared with adjacent antennas.

The center of the antenna is a cast-in-place concrete donut-shaped hub with 12 sleeves for PVC spars which provide the support for the surface and 12 horizontal spars which allow for additional support.  The twelve surface spars are made of 60mm PVC pipe, which are stressed along three support points to approximate a parabolic shape.  Note that an ideal moment-loaded beam actually attains a true parabola.  The three support points (one at the hub, one about 3 meters out radially and one at the rim-end) are at the correct height and angle for the underlying parabola and the PVC essentially acts as a spatial filter.

The shape between the spars is actually a parabola rather than a paraboloid, so the dish surface may more properly be called a ``faceted parabola''.  If only twelve spars were used for the entire surface, simple Ruze losses would be quite high (about 10\% at 150 MHz).  Therefore, at about 2.4 meters another horizontal member is placed, which launches another parabolic spar.  The simple Ruze loss is then less than 1\%, which is much more acceptable.  Fig \ref{fig:elementplan} shows this surface scheme and the panelization (left) and the offsets (in mm) of that scheme from an ideal parabola.  The panel that is ``left out'' is the location of a door to allow access into the hub (via a small removable door and bridge).  The additional intermediate spars also make the panel sizes much more manageable and better secured.  The panel arrangement is also set by a standard width of mesh roll of 1.22m.

\begin{figure}[h!]
\centerline{
\includegraphics[height=6cm]{plots/antPlan.pdf}
\includegraphics[height=6cm]{plots/offset.pdf} 
}
\caption{\small Element panel plan (left) and residual from a true paraboloid (right).
\label{fig:elementplan}}
\end{figure}

\subsubsection{Feed}
\label{sec:feed}
The feed is to be deployed along two separate phases.  Initially, the feed reuses the original PAPER sleeved dipole, however in a new optimized cylindrical feed-cage configuration.  The optimization was based on main beam efficiency, cross-coupling (integrated feed pattern on the dish), standing waves (blockage size), frequency response and polarization match.  The design parameters were size of backplane, height of mast, and height of cylinder.  A range of conical structures were looked at but were ruled out since it was found they introduce additional frequency structure with no real improvement in performance.  The polarization figure-of-merit shown in Fig. \ref{fig:ant_params} is a simplified variant of the fractional power leakage as defined in \cite{Moore_et_al2015}:
\begin{equation}
\xi = \frac{\integral_{-\pi}^{\pi}\left[\sqrt{G(\theta,\phi=0)} - \sqrt{G(\theta,\phi=90)}\right]^2\sin\theta d\theta}
               {\integral_{-\pi}^{\pi}\left[\sqrt{G(\theta,\phi=0)} + \sqrt{G(\theta,\phi=90)}\right]^2\sin\theta d\theta}
\end{equation}
where $G$ is the beam pattern.
The phase 1 design process and outcome is documented in  \cite{feedmemo} and is currently under test on the 19 elements in South Africa. Additional studies are on-going at NRAO and Cambridge University.

The first phase design uses a 172cm diameter backplane, with a 36cm cylinder height.  The dipole is held at 36cm from the backplane and the back of the feed is rigged at 4.9m above the vertex of the dish.  The 10 dB input match for the feed itself spans from about 105 - 200 MHz.  The resultant calculated antenna performance parameters are shown in Fig. \ref{fig:ant_params} and beam-patterns for 110 MHz, 155 MHz and 190 MHz are shown in Fig. \ref{fig:beampatterns}.

\begin{figure}[h!]
\centerline{
\includegraphics[height=0.6\textwidth]{plots/ant_params.pdf}
}
\caption{\small Antenna parameters}
\label{fig:ant_params}
\end{figure}

\begin{figure}[h!]
\centerline{
\includegraphics[height=0.3\textwidth]{plots/pat_110.pdf}
\includegraphics[height=0.3\textwidth]{plots/pat_155.pdf}
\includegraphics[height=0.3\textwidth]{plots/pat_190.pdf}
}
\caption{\small Antenna patterns at 110 MHz, 155 MHz and 190 MHz.}
\label{fig:beampatterns}
\end{figure}

%Additionally, to lessen cross-coupling, a screen fence around the perimeter is also under investigation.  
%Rather than being vertical, it will have a slight angle and potentially some edge treatment to try and lessen diffraction.  
%Tests of the antennas are discussed in a number of papers CITE.  Figure Y shows beam patterns.

The feed is supported by spring-tensioned aramid fiber lines from the three poles via a rigid support to which the feed cage and dipole attach.  The tensioning line goes down centrally to the hub to hold the feed rigidly in place at the correct height.  This also stabilizes the feed in position and allows for strain relief on the coaxial cables.

The low noise amplifier/balun integrates directly to the base of the feed dipole in the same configuration as per PAPER.  The 150m length of 75 $\Omega$ cable used in the initial design also runs down centrally and then over to the post-amplifier module, which is housed near the RFI-shielded container containing the digital electronics.  This will be replaced by the 35m cable/node design, to be discussed.

\subsection{Array Configuration}
\label{sec:arrayConfig}

As a focused experiment, HERA's configuration is optimized for the robust foreground-avoidance approach to EoR power spectrum estimation. 
This requires a densely-packed core of antenna elements, maximizing sensitivity to short baselines---those least contaminated by foreground chromaticity (i.e.\ the wedge). 
Furthermore, the configuration should be highly redundant, both to enable the delay-spectrum strategy of PAPER \citep{parsons_et_al2012a} and redundant baseline calibration \citep{liu_et_al2010,zheng_et_al2014}.
On the other hand, it is also desirable to improve the mapmaking ability of the array for calibration and map-based power spectrum estimation with longer baselines and a densely-sampled $uv$-plane \citep{dillon_et_al2015a}. This tradeoff was closely examined by \citet{dillon_parsons2016} and the ultimate HERA configuration was drawn from one of the designs they considered.

HERA's 320 core elements are arranged in a compact hexagonal grid, split into three displaced segments to cover the $uv$-plane with sub-aperture sampling density (see Fig.~\ref{fig:arrayConfig}). Splitting the core into three sections triples the density of instantaneous sampling of the $uv$-plane, improving HERA's mapmaking ability.Furthermore, the core is supplemented by 30 additional outrigger elements to tile the $uv$-plane with instantaneously complete sub-aperture sampling out to 250$\lambda$ and complete aperture-scale sampling out to 350$\lambda$ (at 150\,MHz). 
This sub-aperture sampling strategy suppresses grating lobes in the instrument's point spread function and provides information for calibrating and correcting direction-dependent antenna responses \citep{dillon_parsons2016}.
Even with the sub-aperture dithering and long baselines, the design is sufficiently redundant to take advantage of redundant baseline calibration technique and robust to the failure of individual elements.
Resulting calibration errors range from $\sim 5\%$ (in the core) to $\sim 10\%$ (for outriggers) of the residual fractional noise per antenna after averaging. 
The split core only increases the gain errors on the core elements by another $\sim 0.1\%$ from a solid hexagonal configuration. HERA's compact and redundant design allows it to fully utilize PAPER's robust approach to calibration and foreground mitigation, but it also provides for excellent imaging capability that can be leveraged to suppress foregrounds and improve access to the 21\,cm reionization signal.

\begin{figure}[h]
	\centering
	\vspace{-5pt}
	\includegraphics[width=1\textwidth,clip]{plots/HERA_Array_Config.pdf}
	\vspace{-25pt}
	\caption{HERA's elements are divided between a 320-element, hexagonally-packed core and 30 outriggers (left). This produces instantaneous
$uv$ coverage at triple the element packing out to 250$\lambda$ at 150\,MHz, supressing grating lobes in the synthesized beam (middle). All 350 elements can be redundantly calibrated 
using the \citet{liu_et_al2010} technique, yielding calibration errors that are a small fraction of the residual noise per antenna (right).}
		\label{fig:arrayConfig}
			\vspace{-10pt}
\end{figure}

\subsection{Analog Signal Path}
HERA's analog signal path from 70--250 MHz with the full performance EoR band of 100-200 MHz, as shown in Fig. \ref{fig:cosmos} and discussed in Sec. \ref{sec:freqs}, emphasizes spectral smoothness and robustness.  Although the low-frequency sky has an intrinsically bright signal level, in the desired ``cold patches'' the sky temperature is only about 30 K at the upper frequency end, so relatively low receiver temperatures are desired.  Note that at the low end of the band (70 MHz), the sky temperature is about 1000 K and across the EoR band it ranges from about 100-400 K.
With receiver temperatures of 30--100 K readily achievable with ambient temperature electronics however
HERA's low-noise amplifiers (LNAs) may be inexpensive, passively cooled components integrated into the
antenna feed.

As part of the delay specification,
HERA's signal path emphasizes careful impedance matching between each component to avoid signal reflections that can worsen foreground leakage.
In particular, impedance matching between free space (377$\Omega$) and coaxial cable (50$\Omega$) is undertaken
in the feed/balun at the front end of the signal chain, where the time constants of inevitable reflections are
minimized.  For a similar reason, HERA houses analog-to-digital converters (ADCs) in nodes near the antenna elements
to limit the total analog path length.  This length is set to be 35m, such that the round-trip delay corresponds to about 0.15 $h$/Mpc, which remains in the wavenumbers  contaminated by the wedge.  The project is also examining the impact of having a range of analog path lengths of $35\pm10$ m.

\subsection{Digital Signal Path}
\label{sec:digital}

\noindent Although correlator development has historically been one of the most 
expensive aspects of building a large radio interferometer, this is no longer the case.
CASPER \Mycitep{parsons_et_al2006}
open-sourced the development of digital signal processing engines for astronomy and
now has world-wide participation,
with over 500 members at 73 institutions, and 
five generations of hardware.
On a modest budget, PAPER applied CASPER technology to develop and deploy new correlators
annually for five years running, each quadrupling the computational capacity of its predecessor.
Led at UC Berkeley's Radio Astronomy Lab (RAL),
HERA efforts continue this incremental development cycle, using a packet-switched
correlator architecture \Mycitep{parsons_et_al2008} that both PAPER and LEDA have
applied to systems employing FPGAs and GPUs \citep{clark_et_al2011}.

In order to meet the 35\,m specification for a maximum analog signal path length, along with a growing need for system scalability, HERA will adopt a node-based architecture for amplification, digitization, channelization, and digital
transmission in the field, building on MWA and Allen Telescope Array heritage. This design is merged with PAPER's clean 
architecture for real-sampling and channelizing the entire analog passband at once, packetizing the data into
10 Gb Ethernet format, and relying on commercial switches to perform the frequency/antenna corner-turn required for FX correlators.  This has led to a new analog-digital-optical board called the Smart Networked Analog-Digital Processor (SNAP), which will become a new modular CASPER device.  Figure \ref{fig:hardware} shows the procession of boards.  This section will step through the digital signal path organized by location:  node, container, processing building.

{\bf Node}. HERA-350 employs RFI-tight node enclosures that each contain the final gain and digitization stages for
signals from 12 antennas, along with power supplies, cooling, sensors, and a small server for monitor/control.  
As part of previous HERA support,
a new Smart Network ADC Processor (SNAP; Fig.~\ref{fig:hardware}) board has been incorporated 
into the Collaboration for Astronomy Signal Processing and Electronics Research (CASPER) suite of hardware and firmware. This inexpensive board was co-designed by UC Berkeley and NRAO to be
both the digitizer and F-engine in HERA's FX correlator architecture.
Each SNAP board 
digitizes and channelizes a 0--250 MHz band for 6 input signals (3 antennas, dual-polarization).
A 200-MHz band of selectable channels is transmitted over optical fiber
to a central container (see below), and on to the correlator.  Laboratory
tests have successfully demonstrated robust transmission from the SNAP board, through 10 km of fiber optic cable, 
into a 10 GbE switch using inexpensive, commercially available optical transceivers.
Activities under this proposal include 
testing and extending FPGA firmware,
integrating and testing all components in the node subsystem, and providing a monitor/control
access interface.

\begin{figure}[h]
\centering
%\includegraphics[width=2.2in]{plots/ROACH2.png}
\includegraphics[width=1.0\textwidth]{plots/casper_boards.png}
%\includegraphics[height=1.7in]{plots/snap_board.png}
\caption{
Six generations of CASPER digital signal processing (left to right) culminating in the SNAP board (right, along with the long-haul fiber-link test setup), which was developed and tested under the previous HERA MSIP award.
By preserving its toolflow, signal processing libraries, and interface code between hardware generations,
CASPER combines with a scalable correlator architecture to enable the PAPER correlator to be 
easily upgraded for HERA \Mycitep{parsons_et_al2006,parsons_et_al2008}.
%
}\label{fig:hardware}
\end{figure}

{\bf Central Container}.
HERA's central container houses two significant subsystems adjacent to the array.  The first is a timing subsystem
that maintains a GPS-disciplined oscillator and distributes timing
signals (the sampling clock and 1 pulse-per-second synchronization) to the nodes.  The second
subsystem is a passive fiber optic patch panel that couples
the optical network from the nodes into the 192-filament optical fiber bundle 
that connects to the KAPB.

{\bf Karoo Array Processing Building (KAPB)}.
The KAPB houses the switch and CPUs
that 
complete the HERA correlator system.  A fiber optic bundle enters the KAPB and patches
into local fiber optic cables. These cables terminate in optical transceivers which plug into a 240-port 10 GbE switch.
Such switches, while large, are readily available commercially today.  Also connected to
this switch are 30 servers, each hosting two dual-GPU graphics cards and two dual
10 GbE network interface cards, which implement the cross-multiplication (X-Engine) component
of the correlator during observations.  This estimate for the number of X-Engine servers
is extrapolated from current GPU servers deployed on PAPER, assuming no improvement in bus
speeds for transferring data into the GPU cores, but assuming that the computational
capacity of GPU cores doubles once according to Moore's Law prior to purchasing
these servers in year three of the project.
Output data from the correlator are written to the data storage system described
in the following section.

\subsection{Data Processing and Management}
\label{sec:data}
\noindent The HERA correlator generates $\sim4$~TB of raw data per 12-hour observing day.
HERA's data management system archives this on-site in a 1.5 PB storage array, calibrates it in real-time
on the basis of redundancy, compresses it by a factor of 20, and transfers it to our computing facility at NRAO.
This real-time pipeline, based on a similar system for PAPER, is mature, computationally manageable, and 
agnostic with respect to observing parameters, making it a safe and robust first step in data processing.  For a discussion
of the PAPER pipeline, see
\cite{zheng_et_al2014,parsons_etal2014,ali_et_al2015}.
The Internet link between the Karoo and data centers in the U.S.\ has been extensively tested, with assistance from SKA-SA, TENET
(South African to London link provider) and GEANT (London to U.S.\ provider).   
Transfer speeds are sufficient for streaming compressed data to the U.S.\ via the Internet. Portable storage is also
provisioned as a fallback transfer method.

NRAO hosts the archiving, indexing, and processing of HERA data products. 
Project data and public data products are served via the existing 
NRAO archive at the Domenici Science Operations Center %(DSOC) 
in Socorro, New Mexico, where it will be accessible through computing accounts and public web-based
archive searches.
NRAO also supports
an attached 30-node processing cluster 
coupled to 
scratch storage
for routine data inspection and lightweight analysis tasks.
For the bulk application of processing-intensive data pipelines to HERA data,
NRAO manages the dynamic acquisition of computing from
external HPC and cloud-computing facilities, including the Amazon spot market (\url{aws.amazon.com/ec2/spot}) and XSEDE (\url{www.xsede.com}) . 
Combined with legacy computing resources at UPenn and MIT, this plan leverages NRAO's
expertise in managing and acquiring computing resources while providing distributed
permanent resources for HERA's basic processing needs.

%This effort leverages existing infrastructure at NRAO, with support 
%for the expansion of the data storage and for upgrading to a 30-node computing cluster.  This cluster supports the bulk of the analysis by HERA collaborators (see \S\ref{sec:analysis}) that requires access to the full set of HERA observations.

\subsection{Analysis Pipelines} 
\label{sec:software}

\noindent HERA builds on the rich legacy of PAPER and MWA software and database systems 
developed for field operations, data analysis, and simulation.  Examples range from
the strictly versioned and unit-tested packages for field-deployed systems (e.g. the correlator, real-time processing, and
monitor/control systems) to loose collections of scripts written for exploratory analysis.
Software packages that support HERA analysis are open source, publicly 
hosted\footnote{Including \url{github.com/AaronParsons/aipy}, \url{github.com/JeffZhen/omnical}, 
\url{github.com/MiguelFMorales/FHD}, \url{github.com/MiguelFMorales/eppsilon}, \url{github.com/jpober/21cmSense},
and \url{github.com/Nithyanandan/PRISim}.}, revision controlled, and unit-tested.  
These include OMNICAL, a complete package for redundant baseline calibration;
Astronomical Interferometry in Python (AIPY), a set of
tools and file-format interfaces for reading visibilities, calibrating,
rephasing, imaging, and deconvolution; Fast Holographic Deconvolution
(FHD), a purpose-built tool for imaging, calibration, and foreground
forward-modeling and subtraction \citep{sullivan_et_al2012}; Precision Radio Interferometry Simulator 
(PRISim), a package for accurately simulating wide-field interferometric observations;
21cmFAST \citep{mesinger_et_al2011}, a fast, semi-numerical 21~cm signal simulator,
and 21cmSense \citep{pober_et_al2014}, a tool for forecasting power spectrum sensitivity.
Other project code is aggregated and revision
controlled in a public repository with separate sandboxes for each developer to ensure that
HERA members have up-to-date copies of all project code to facilitate sharing and debugging.
Such code includes the PAPER pipeline for foreground filtering and estimating power spectra from
visibility data,
the MWA power spectrum analysis codes---\eppsilon\ \citep{hazelton_et_al2016} and the 
empirical covariance technique of \citet{dillon_et_al2015}---as well as
machine-learning-based source finding, verification, and removal tools \citep{caroll_et_al2016,jacobs_et_al2016,beardsley_et_al2016}.

New software development will focus on integrating and improving the MWA and PAPER power spectrum and foreground removal pipelines,
developing a monitor and control software interface and database for recording instrument metadata,
extending, with support from Scuola Normale Superiore, semi-analytic and numerical models 
of the 21~cm signal for robust parameter estimation, and developing
machine learning interpolations of simulations for joint Monte Carlo estimation of cosmological and astrophysical parameters.

\section{Schedule and Status}
\label{sec:status}

\noindent On-site construction will proceed in stages, starting from the
initial 37 elements that will be in place by project start.  These will be used
in conjunction with the thoroughly characterized extant PAPER signal path and
processing hardware for a very low risk initiation of scientific observations
from the very beginning.  Additionally, the 19 HERA elements currently in
place will also have had an extensive period of commissioning
and characterization.  As elements 38 to 128 are installed in year 1, they can
immediately be placed within the array and be used for observing.

During this period, infrastructure for the new node-based system is installed
and tested with the first elements beyond 128.  After the HERA-128 observing
season, the full array is transitioned to use HERA's new hardware infrastructure.  In this same time frame, the existing PAPER processing
container is moved to the edge of the array to house the timing sub-system
and fiber optic splice cabinet while the correlator will move to the KAPB.
Also, the data processing activity at Penn is transitioned to NRAO.
Summarizing, the hardware deployment phases are in four categories:
\begin{enumerate}
\item Element construction:  linear process (years 1, 2 \& 3)
\item Infrastructure:  (a) node power and fiber reticulation (years 1 \& 2), (b) container relocation (year 1), (c) node installation (years 2 \& 3)
\item Signal path:  convert to node-based system after 128 elements (years 2 \& 3)
\item Processing:  (a) data processor/server to NRAO (year 1), (b) X-engine \& real-time processor/storage to KAPB (year 2)
\end{enumerate}

\begin{figure}[h]
	\vspace{-7pt}
	\begin{tabular}{ll}
	\begin{minipage}{5in}
		\includegraphics[width=4.9in]{plots/timeline_short.pdf}
		\end{minipage} & \hspace{-.15in}
	\begin{minipage}{1.35in}
\captionsetup{justification=raggedright,
singlelinecheck=false
}
	\caption{Timeline of HERA construction, analysis development, observation, and scientific output.}
	\label{fig:timeline}
	\end{minipage}
	\end{tabular}
	\vspace{-8pt}
\end{figure}

The optimal EoR observing window is from September to April, with power spectrum limit results appearing about one year later.  Concurrent technique development and deployment will be on-going. Additional information is in the project management plan.
\begin{itemize}[leftmargin=0.7in]
\item[Year 1:] Observe with H37.  Real-time data pipeline, delay-space power spectrum (DSPS) pipeline, FHD pipeline.  21\,cm framework for incorporating with other probes. Construct H128.
\item[Year 2:] Observe with H128.  Real-time calibration pipeline.  DSPS/FHD/global sky model integration. Snapshot imaging pipeline.  EoR parameter estimation development.  H37 results.  Construct H240.  {\em Data products:  power spectrum, Stokes I maps.}
\item[Year 3:] Observe with H240.  Foreground-filtered imaging pipeline.  EoR parameter estimation development.   H128 results. Construct H350.  {\em Data products:  power spectrum, Stokes IQUV maps, foreground image cube.}
\item[Year 4:] Observe with H350.  EoR parameter estimates.  H240 results.  {\em Data products:  power spectrum, global sky model IQUV, snapshots, foreground-filtered image cube.}
%\item[Year 5:] Observe with H350.  H350 results.  {\em Data products:  power spectrum, global sky model IQUV, snapshots, foreground-filtered image cube.}
\end{itemize}

\section{Conclusion}
\label{sec:conclusion}

\noindent With first-generation instruments, the HERA team has led a revolution in our understanding of how to best perform 21\,cm cosmology measurements.
In the past three years, we have developed the EoR window paradigm for isolating foreground systematics, implemented novel
calibration and power spectrum analysis pipelines, made precision measurements of astrophysical foregrounds, and published deep power 
spectrum limits that constrain heating in the early universe.  HERA's design, now validated in the field with MSIP seed funding, 
is the result of these advances.

HERA is ``the right instrument at the right time.'' With first generation experiments approaching the limits of their sensitivity, it is now critical to build upon our new knowledge and experience to advance beyond them. HERA's large collecting area and optimized design enable precise constraints of EoR astrophysics and a broad range of high-impact secondary science, all before SKA-low becomes operational.
With CHAMP, HERA expands CAMPARE's proven methods and recruitment network to bring more students from underrepresented groups into cutting-edge astronomy research and support them with year-round mentoring.
By bringing the U.S.\ teams from PAPER, the MWA, MITEoR, and EDGES together with strategic international partners,
the HERA collaboration has consolidated theoretical and experimental leadership in the field, 
and is equipped to deliver the proposed science. By building HERA, we can provide the community with unprecedented insights into this key new area of cosmology.

\clearpage
\setcounter{page}{1}
\thispagestyle{empty}
%\bibliographystyle{apj}
%\bibliographystyle{hapj}
\nocite{Beardsley:thesis}
\nocite{kolopanis_et_al2016}
\bibliographystyle{jponew}
%\bibliographystyle{unsrt}
{\small \bibliography{biblio}}

\end{document}